---
title: "simple_doc"
author: "Fangyi Chen"
date: "2023-09-19"
output: github_document
---

Let's import the 'FAS_litters.csv'


```{r}
library(tidyverse)
library(readxl)
```


```{r}
litters_df = read_csv(file="data_import_examples/FAS_litters.csv")

litters_df = janitor::clean_names(litters_df)
```

Import dataset using absolute path
```{r}
litters_df = read_csv(file="/Users/madgeniuschen/Desktop/data_science/data_wrangling/data_import_examples/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

**Learning Assessment**: Make sure you are able to load the FAS_pups.csv dataset. Use both absolute and relative paths. Quit R Studio and move the directory containing your project, data, and R Markdown document. Repeat the previous data import process; do both absolute and relative paths still work?


```{r}
pups_df = read_csv(file="data_import_examples/FAS_pups.csv")

pups_df = read_csv(file="/Users/madgeniuschen/Desktop/data_science/data_wrangling/data_import_examples/FAS_pups.csv")

pups_df  = janitor::clean_names(pups_df )
```


# Look at the data 

```{r}
head(litters_df)
tail(litters_df)

```
```{r, eval=FALSE}
view(litters_df)
```


Look at data summary 
```{r}
str(litters_df)
```

```{r}
skimr::skim(litters_df)
```

## Data Parsing 
```{r}
litters_df = read_csv(file="data_import_examples/FAS_litters.csv",
                      skip=10, col_names = FALSE)


```

Look at NA values

```{r}
litters_df = read_csv(file="data_import_examples/FAS_litters.csv",
                      na = c("NA", 19))  # can treat 19 or other customized vals as missing
litters_df
```

columns types
```{r}
litters_df = read_csv(file="data_import_examples/FAS_litters.csv", 
                      col_types = cols(
                        Group = col_factor()
                      ))
litters_df

```

# Other file types

import a xlsx file first
```{r}
mlb_df = read_excel("data_import_examples/mlb11.xlsx")
```


```{r}
library(haven)
pulse_data = read_sas("data_import_examples/public_pulse_data.sas7bdat")
head(pulse_data, 5)
```



```{r}
litters_df$Group
```


```{r}
litters_df = read_csv(file="data_import_examples/FAS_litters.csv")

litters_df = janitor::clean_names(litters_df)
write_csv(litters_df, "data_import_examples/clean_litter.csv")
```

